---
title: "Final Exam"
subtitle: "STAT 201"
author: "Anjali Chepyala"

format:
  html:
    toc: true
    embed-resources: true

execute:
  message: false
  warning: false
---

# Instructions

1. **Write your name** at the top! Major deductions if not.

2. Write your code in the *Code* cells of the **template provided** to write solutions for the assignment. **Do not open a new qmd**, and work from scratch. Ensure that the solution is written neatly enough to understand and grade.

3. This exam is open note open book. You are NOT permitted to use AI or each other.

4. There are point deductions if your code is not clean and organized.

5. Highly inefficient code will be deducted for because it demonstrates lack of understanding. 

6. Each question should print your answer. But don't print excessive/random information.

7. Render as an HTML and **submit BOTH the .qmd and HTML file**.

I cannot grade what I cannot see! Failure to render your document after completing all questions can lead to major deductions.


## Question 1 (*2 points*)

Continually **generate** a random integer between 1 and 10 until the sum is greater than or equal to 30. 

Print the random integer that is generated.

Once the program is **complete** print the following statement:

"You generated {} numbers that sum to {}." 

Note:

- the first {} is the count of the number of terms entered 
- the second {} is the sum of all the terms
    
Run your program until it ends.

```{r}
x <- 0
countloop <- 0
for (i in 1:30){
  while (x <=30){
    v <- sample(1:10, 1, replace=TRUE)
    x <- x + v
    countloop <- countloop + 1
  }
}
print(sprintf("You generated %d numbers that sum to %d", countloop, x))
```

## Question 2 (*2 points*)

Write a function that calculates BMI. 

The function should have inputs for height (in inches) and weight (in pounds).

BMI =  $703 * \frac{weight}{height^2}$

BMI **categories** are defined as follows:

- Underweight: < 18.5
- Normal: 18.5 <= BMI < 25
- Overweight: 25 <= BMI < 30
- Obese = BMI of 30 or greater

The function should return 2 values: the BMI (numeric value) and category ("Underweight", "Normal", "Overweight", "Obese").

**After** your function is defined, complete the following tasks:

- Run your function using any height and weight of your choice
- Store the output of the function as a variable(s)
- Print the following statement **using your stored variable** with BMI rounded to 1 decimal place:

                   "Your BMI is {} with a value of {}." 
          Example: "Your BMI is Normal with a value of 22.5."
                
```{r}
BMI_calculate <- function(height, weight){
  BMI <- 703 * (weight / height^2)
  
  if (BMI >= 30) {
    print(sprintf("Your BMI is Obese with a value of %f", BMI))
  } else if (BMI >= 25) {
    print(sprintf("Your BMI is Overweight with a value of %f", BMI))
  } else if (BMI >= 18.5) {
    print(sprintf("Your BMI is Normal with a value of %f", BMI))
  } else {
    print(sprintf("Your BMI is Underweight with a value of %f", BMI))
  }
}

BMI_calculate(64, 115)
```
         
                
## Question 3 (*5 points*)

### a

Create 3 vectors named `vec1`, `vec2`, `vec3`: 

- `vec1` should contain the numbers: 5, 8, -2, 0, 11
- `vec2` should contain the numbers: -7, 3, 12, 6, -1
- `vec3` should be a `seq()` of 1 to 13 by 3

```{r}
vec1 <- c(5, 8, -2, 0, 11)
vec2 <- c(-7, 3, 12, 6, -1)
vec3<- seq(1, 13, 3)
```


### b

Create a matrix where `vec1` is the first column, `vec2` is the second column, and `vec3` is the third column.

Print the matrix.

```{r}
matex <- matrix(vec1, byrow=FALSE)
matex <- cbind(matex, vec2)
matex <- cbind(matex, vec3)
colnames(matex) <- c("vec1", "vec2", "vec3")
print(matex)
```


### c

Use an appropriate apply function to calculate the maximum value of each column of the matrix. Store the output as `max_col` then print the results.

```{r}
maxnum <- c()
for (i in 1:ncol(matex))
{
  maxnum[i]= max(matex[,i])
  print(maxnum[i])
}
    
  
    
```


### d

Create a list object where each element is one of the vectors (`vec1`, `vec2`, `vec3`) from part a. 

Print the list.

```{r}
listob = list(vec1, vec2, vec3)
print(listob)
```


### e

Use an appropriate apply function to determine if the `sum` of each list element is even. The output should be stored as a vector and contain either TRUE or FALSE (depending if the sum was even or odd). 

Print the results.

Note: if done correctly this will print: TRUE FALSE FALSE

```{r}
sumeven <- function(x){
  fulllist <- unlist(listob)
  vec1sum <- sum(vec1)
  vec2sum <- sum(vec2)
  vec3sum <- sum(vec3)
  if (vec1sum %%2 == 0){
    print("TRUE")
  } else{
    print("FALSE")
  }
  if (vec2sum %%2 == 0){
    print("TRUE")
  } else{
    print("FALSE")
  }
  if (vec3sum %%2 == 0){
    print("TRUE")
  } else{
    print("FALSE")
  }
}
sumeven(listob)
    
    
  
```


## Question 4 (*3 points*)

Consider the `fruit_info` object defined below which contains fruits, their corresponding color, and their weight in grams.

```{r}
fruit_info = data.frame(
  fruit = c("apple", "banana", "grape", "strawberry",  "watermelon", 
            "lime", "lemon", "cherry", "cranberry", "orange", "pineapple", 
            "avocado", "blueberry"),
  color = c("red", "yellow", "green", "red",  "red", 
            "green", "yellow", "red", "red", "orange", "yellow",
            "green", "blue"),
  weight = c(80, 108, 5, 13, 6450, 67, 58, 11, 2, 120,
             900, 215, 3)
  )
```


### a

Use slicing to create a data frame of only fruits that are red.

Store this new data frame as `fruits_red` and print the object.

```{r}
fruits_red <- fruit_info[fruit_info$color == "red", ]
print(fruits_red)
```


### b

Use an appropriate built-in function (no loop) to print the "name" of the `fruit` that has the heaviest weight. 

Only the name of the fruit should be output/printed.

```{r}
weightweight <- fruit_info[fruit_info$weight, ]
print(weightweight)
heavy <- fruit_info[fruit_info$weight == 6450, ]
print(heavy)
```


### c

If a fruit is over 100 grams it is "heavy", otherwise it is "light".

Using any method of your choice, print the following statement indicating whether each fruit is heavy or light.

Your first two printed lines will read as follows:

  - "The apple is light."
  - "The banana is heavy."
  - etc.

```{r}
for (j in 1:ncol(fruit_info)) {
  # your code here, for example:
  print(paste("Column", j, "name is", colnames(fruit_info)[j]))
}
```


